### 2015-10-22 ManuscriptMarkDown.Rmd
Initial commit

---
title: "Monte Carlo Tests For Species Co-occurrence"
author: "Nick Gotelli"
date: "October 21, 2015"
output: 
  html_document: 
    highlight: haddock
    keep_md: yes
    number_sections: yes
    theme: cerulean
---

## Introduction
## 2 x 2 table of species associations
## Binary presence absence matrix
## Null model procedure
## Fixed rows
## Fixed columns
## Fixed-fixed curve ball algorithm
## Other approaches